\propertitle

### 2019-06-17 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-17T11:00:59+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
% https://tex.stackexchange.com/a/33547/97039

@@

Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
maximum-entropy method. In short, it amounts to a pair of prescriptions:
choose the distribution having maximum Shannon entropy among those
satisfying specific convex constraints, such as fixed expectation values;
and judge these expectation values to be equal to measured averages. We add
two remarks about this method that are seldom made in the literature.
First, the \enquote{distribution} given by this method is the zeroth-order
approximation of four different distributions:
\begin{itemize}
\item the \emph{probability} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
\item the \emph{probability} distribution for the value of the total
  activity at a \emph{new} time bin,
\item the most probable \emph{time-frequency} distribution of the total
  activity across the \emph{recorded} bins,
\item the most probable \emph{time-frequency} distribution of the total
  activity in a very long run of \emph{new} time bins.
\end{itemize}
The maximum-entropy distribution is thus an approximate answer to four
conceptually different inferences. Note that although the four
distributions above numerically coincide in this approximation, they differ
in higher-order approximations. We discuss their differences in \sect***.
Second, the maximum-entropy method based on the Shannon entropy implicitly
makes specific assumptions about the probability of the long-run frequency
distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We
discuss these assumptions in \sect***.

@@

These two steps are more straightforward if instead of power moments we use
\emph{normalized factorial moments}. The $m$th normalized factorial moment
of the probability distribution for the activity of the recorded neurons is
defined as the expectation
\begin{equation}
  \E\condt[\Bigg]{
    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
  \qquad 1\le m \le n
\end{equation}
where $\pf(\ya \| \yH)$ is the probability for activity $\ya$ given
assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
new bin. This moment can be interpreted as the expectation of the number of
distinct $m$-tuples of simultaneously spiking neurons (within a bin's time
width), normalized by the number of distinct $m$-tuples. For example, with
$m=2$, if $na=4$ neurons spike in a network of $n=5$, we have
$\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons, and the
total number of distinct pairs is $\tbinom{5}{2}=10$. The normalized number
of pairs is therefore $6/10$. Note that the first $m$ factorial moments
linearly determine the first $m$ moments and vice versa, so fixing the ones
is equivalent to fixing the others. Now, since the recorded neurons are a
sample from the full network of $N$ neurons, we have this extremely
convenient property: \emph{The first $n$ normalized factorial moments for
  the sample and for the full network are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad m \le n.
\end{equation}

@@

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-17T16:26:01+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
% https://tex.stackexchange.com/a/33547/97039
\newcommand*{\fm}{f}
\newcommand*{\Fm}{F}

@@

Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
maximum-entropy method. In short, it amounts to a pair of prescriptions:
choose the distribution having maximum Shannon entropy among those
satisfying specific convex constraints, such as fixed expectation values;
and judge these expectation values to be equal to measured averages. We add
two remarks about this method that are seldom made in the literature.
First, the \enquote{distribution} given by this method is the zeroth-order
approximation of four different distributions:
\begin{itemize}
\item the most probable \emph{time-frequency} distribution of the total
  activity across the \emph{recorded} bins,
\item the most probable \emph{time-frequency} distribution of the total
  activity in a very long run of \emph{new} time bins,
\item the \emph{probability} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
\item the \emph{probability} distribution for the value of the total
  activity at a \emph{new} time bin.
\end{itemize}
The maximum-entropy distribution is thus an approximate answer to four
conceptually different inferences. Note that although the four
distributions above numerically coincide in this approximation, they differ
in higher-order approximations. We discuss their differences in \sect***.
The maximum-entropy distribution approximates best the first of the four
distributions, so that's the one we'll mainly have in mind, and we'll
denote it by $\Fm(\yA)$.***
Second, the maximum-entropy method based on the Shannon entropy implicitly
makes specific assumptions about the probability of the long-run frequency
distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We
discuss these assumptions in \sect***.

@@

These two steps are more straightforward if instead of power moments we use
\emph{normalized factorial moments}. The $m$th normalized factorial moment
of the probability distribution for the activity of the recorded neurons is
defined as the expectation
\begin{equation}
  \E\condt[\Bigg]{
    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
  \qquad 1\le m \le n
\end{equation}
where $\pf(\ya \| \yH)$ is the probability for activity $\ya$ given
assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
new bin. This moment can be interpreted as the expectation of the number of
distinct $m$-tuples of simultaneously spiking neurons (within a bin's time
width), normalized by the number of distinct $m$-tuples. For example, with
$m=2$, if $na=4$ neurons spike in a network of $n=5$, we have
$\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons, and the
total number of distinct pairs is $\tbinom{5}{2}=10$. The normalized number
of spiking pairs is therefore $6/10$. Note that the first $m$ factorial
moments linearly determine the first $m$ power moments and vice versa,
because $\tbinom{n\ya}{m}$ is a polynomial in $\ya$ of degree $m$; so
fixing the ones is equivalent to fixing the others. Now we have this
extremely convenient property: \emph{the first $n$ normalized factorial
  moments for a sample and for the full network are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad m \le n.
\end{equation}

@@

We apply the approach just described to two data sets publicly available in
the literature:

The first data set, from \textcite[rat 14147]{stensolaetal2012}, consists
of $n=65$ neurons (27 of which classified as grid cells) from rat Medial
Entorhinal Cortex, recorded for about 20 minutes. Their spikes are
binned ainto $T=417\,641$ bins of $3\;\textrm{ms}$ width.

The second data set, from \textcite{rostamietal2016_r2017}, consists of
$n=159$ neurons from macaque Motor Cortex, recorded for about 15 minutes.
Their spikes are binned ainto $T=300\,394$ bins of $3\;\textrm{ms}$ width.

\texttt{data\_RS.npy}
